---
title: "Bike Sharing Demand"
author: "Minghao Liu"
date: "11/11/2019"
output: html_document
---


# Bike Sharing Demand Kaggle challenge

## Coefficients:

## datetime - hourly date + timestamp
## season - 
###  1 = spring, 2 = summer, 3 = fall, 4 = winter
## holiday - whether the day is considered a holiday
## workingday - whether the day is neither a weekend nor holiday
## weather -
###  1: Clear, Few clouds, Partly cloudy, Partly cloudy
###  2: Mist + Cloudy, Mist + Broken clouds, Mist + Few clouds, Mist
###  3: Light Snow, Light Rain + Thunderstorm + Scattered clouds, Light Rain + Scattered clouds
###  4: Heavy Rain + Ice Pallets + Thunderstorm + Mist, Snow + Fog
## temp - temperature in Celsius
## atemp - "feels like" temperature in Celsius
## humidity - relative humidity
## windspeed - wind speed
## casual - number of non-registered user rentals initiated
## registered - number of registered user rentals initiated
## count - number of total rentals


# To load data and packages, set theme
```{r}

df <- read.csv('/Users/AncelotLiu/Desktop/R-Course-HTML-Notes/R-for-Data-Science-and-Machine-Learning/Training\ Exercises/Machine\ Learning\ Projects/CSV\ files\ for\ ML\ Projects/bikeshare.csv')
class(df)

head(df)

library(ggplot2)
library(ggthemes)

theme_set(theme_economist_white())

```

# Create a scatter plot of count vs temp. Set a good alpha value.
```{r}


ggplot(df, aes(temp, count, color = temp)) + geom_point(alpha = 0.5)


```

# Plot count versus datetime as a scatterplot with a color gradient based on temperature. You'll need to convert the datetime column into POSIXct before plotting.
```{r}
df$datetime <- as.POSIXct(df$datetime)

ggplot(df, aes(datetime, count, color = temp)) + geom_point(alpha = 0.5)
```

# The correlation between temp and count
```{r}


cor(df[,c('temp','count')])

```

# Season data: Create a boxplot, with the y axis indicating count and the x axis begin a box for each season.
```{r Season Data}


ggplot(df,aes(x=factor(season),y=count, color = factor(season))) + geom_boxplot()

```

# Create an "hour" column that takes the hour from the datetime column.
```{r hour column}

hour <- format(df$datetime, '%H')

df <- cbind(df, hour)

head(df)
```

# Now create a scatterplot of count versus hour, with color scale based on temp. Only use bike data where workingday==1.
```{r count vs hour, working day}

ggplot(df[df$workingday == 1,], aes(x=hour,y=count,color = temp)) + geom_point(position=position_jitter(w=1, h=0)) + scale_color_gradientn(colors=c('black','blue','green','yellow','red'))

```

```{r count vs hour, non working day}

ggplot(df[df$workingday == 0,], aes(x=hour,y=count,color = temp)) + geom_point(position=position_jitter(w=1, h=0)) + scale_color_gradientn(colors=c('black','blue','green','yellow','red'))

```

## Notice that working days have peak activity during the morning (~8am) and right after work gets out (~5pm), with some lunchtime activity. While the non-work days have a steady rise and fall for the afternoon